If you see something, then say something to others: Complex contagions and the socially-contingent correction of misinformation

### 2018-02-28 manuscript_1.0.tex
Revised networks and codes, working on SEIR dynamic codes
- draft updates until EI progression module

% declare documents and packages
\documentclass[man, 12pt, a4paper, noextraspace]{apa6}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[american]{babel}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{microtype}
\usepackage{caption}
\usepackage{subcaption}

\usepackage{graphicx}
\usepackage{url}
\usepackage{times}
\renewcommand{\footnotesize}{\small} % use 11pt for footnotes
\usepackage[doublespacing]{setspace} % single space for footnotes
\setlength{\skip\footins}{1.5pc plus 1pt} % some more space between footnote and main body
% this command put all footnotes to endnote at the end of the paper
%\usepackage{endnotes}
%\let\footnote\endnote

@@

\note{Word count: 5339 words \\ %\quickwordcount \\
    Draft date: \today \\
      \addvspace{.25in} 
      \textbf{Draft in progress. Please do not cite without permission.} \\
      Please direct any questions and inquiries to \href{mailto:hyunjin.song@univie.ac.at}{hyunjin.song@univie.ac.at}}

\begin{document}
\setcounter{page}{0}
\maketitle
  Citizens across the worlds are experiencing major changes in their news environments with the development of digital media. One of the most dramatic changes in the news environment in recent decades involves the role social networking sites (SNS) such as Facebook and Twitter play as a primary source of news outlets. Not only citizens' news consumptions are increasingly driven by such online sources \parencite{shearer2017news}, but it also appears that citizens themselves are actively participating in news dissemination on those platforms by sharing news contents with their peers \parencite[e.g.,][]{shearer2017news, lee2017people}.

@@

      Another line of studies based on a dual process theory of human cognitive processing and memory suggests that attitudinally-congruent misinformation creates automatic and strong affective responses -- therefore automatically and effortlessly activated in one's memory -- whereas attitudinally incongruent correction messages rarely produce such responses. Due to such asymmetrical nature, people have to rely on more deliberative, strategic search processes (which require significant cognitive resources) to recall attitudinally inconsistent correction messages and incorporate them into relevant judgments \parencite{thorson_2016, Lewandowsky_2012PSPI}. Also, since misinformation tends to form a coherent mental model based on one's partisan schema and stereotypes \parencite[e.g.,][]{garrett2013undermining}, people tend to fill any gaps caused by corrections (that invalidate some parts of the existing mental model) with flawed but attitudinally congruent misinformation that is still readily accessible in their memory \parencite{Lewandowsky_2012PSPI}. Studies also find that this effect is much more likely when correction messages do not update the initial mental model that justifies misinformation \parencite{Chan_debunking_meta_2017}, when the perceived veracity of initial misinformation is high \parencite[due to fluency bias in one's cognitive processing:][]{Lewandowsky_2012PSPI}, or when individuals can generate counter-arguing reasons in support for initial misperceptions \parencite{garrett2013undermining, Chan_debunking_meta_2017}. Most importantly, due to limitations of strategic memory search processes (i.e., it requires more effortful processing), people may still rely on negated misinformation in subsequent reasoning \emph{even when they remember such information is factually incorrect} \parencite{Lewandowsky_2012PSPI}. Therefore, even in the face of seemingly effective corrections, the effect of misperceptions lingers and continue to exert influence \parencite{thorson_2016}.

  Under certain situations, it seems that citizens \emph{indeed} can adhere factual information based on correction messages despite their perpetual partisan bias \parencite[e.g.,][]{Wood2018, nyhan2017taking}. Yet as \citeauthor{margolin2017} (\citeyear{margolin2017}) note, it appears that such effects often require special \emph{social context}. This observation is indeed much warranted, as most of the previous studies concerning misinformation and the effect of fact-checking messages are conducted in an experimental context with a single-shot, \emph{asocial} correction message from media professionals and fact-checking organizations \parencite[e.g.,][]{nyhan2010corrections,garrett2013undermining,weeks2015emotions}. Much of the literature on partisan selective exposure and political discussion networks already point that social contexts of which an individual is exposed to counter-attitudinal messages may have a powerful consequence on how such messages are interpreted and processed \parencite{messing2014selective, levitan2008resistance}. There is also a suggestive evidence that fact-checking and corrective messages from one's peers in their social networks -- what we would call a \enquote{social correction} -- are more likely to, if not equally, be effective in reducing misperceptions \parencite[e.g.,][]{margolin2017, bode2017see}. In what follow, we review several theoretical accounts of such \emph{socially-based} correction messages on misinformation and fake news.

@@

  If the diffusion of (mis)information and adoption of corrective messages may non-trivially dependent upon a structure of a given network and its attitudinal composition, then how a typical (online) social network is structured in terms of its topological features, and how pervasive is \enquote{attitudinal} homophily, as \emph{the} critical factor determining attitudinal compositions, on such a social network? How such structural features affect the overall diffusion dynamics empirically? This is indeed important questions to ask, since the flow of information (either misinformation or its correction) and its adoption are ultimately structured by how individuals are connected with each other in a given network. However, it is surprisingly difficult to establish convincing evidence of the impact of structural properties of a given network and its attitudinal composition in complex contagion dynamics using purely observational and experimental approaches.

@@

    Similarly to attitudinal composition of citizen's social networks, the question of how exactly the structure of citizens' social networks exhibits certain topological properties have attracted considerable interests among scholars. Prior observations on this topic suggests that large-scale (online) social networks tend to exhibit \enquote{small-world} like properties \parencite{kumar2010structure, ugander2011anatomy}. A small world is characterized by its typical structure where most nodes can be reached from every other nodes by a small number of steps compared to a pure random graph of the same size. This is due to the fact that a small minority of \enquote{hub} nodes posses a disproportionate number of links, providing a global bridge between smaller, more strongly interconnected local clusters \parencite{barabasi2004linked}.

    This structure has traditionally been known for information to be spread globally through such hubs \parencite[e.g.,][]{Bakshy_2012}, following \citeauthor{granovetter1977strength}'s (\citeyear{granovetter1977strength}) seminal account of strength of weak ties. However, emerging evidence suggests that such a small-world structure may hinder a given diffusion process to globally saturate the network, since such hubs can serve as a \enquote{bottleneck} of a contagion process, therefore diffusions are stay localized and compartmentalized \parencite[e.g.,][]{banos2013diffusion, centola2007complex, zhao2010weak}. Similarly, while the existence of tightly-knit neighborhoods in a typical small-world network supports a complex-type behavioral propagation \parencite[i.e., an innovation that runs counter to prevalent norms and values:][]{Centola2010Sience}, the existence of hubs may not favor complex contagion to be globally propagated into a given network. This is due to the fact that such hubs require a much higher number of active neighbors to meet the threshold of adopting a controversial innovation (i.e., complex contagions), and even if such hubs are activated, they alone cannot further activate their immediate neighbors to adopt such  innovations in the absence of other local nodes that provide multiple independent reinforcements \parencite{Centola2007449, zhao2010weak, gonzalez2017decoding}. Therefore, whether topological properties of a given network actually hinder or facilitate a contagion process may critically dependent upon specific thresholds values for complex contagions and its interaction with the structure of network \parencite[e.g.,][]{Centola2007449}, which is often hard to be tractable empirically.

    Moreover, especially for the case of misinformation and its corrections, a simple contagion (of misinformation) and a complex contagion (of correction messages) may simultaneously affect a local-level attitudinal composition of one's neighbors \parencite[i.e., those who endorse and believe misinformation vs. those who do not: e.g.,][]{campbell2013complex}. Last but not least, all of aforementioned factors are endogenously determined through prior states of networks over time in evolving dynamic network. As a consequence, a proper identification regarding the exact impacts of its attitudinal composition and structural properties of an given network often necessitates systematic comparisons between the observed network and a number of plausible counterfactuals. Doing so further requires (a) the ability to compared different thresholds values for complex contagions, and (b) the ability to independently manipulate topological structures of a social network \parencite[e.g.,][]{Centola2010Sience, Centola2007449, gonzalez2017decoding}. For this reason, a rigorous empirical test of such factors is often practically impossible for observational studies despite its scientific and practical importance.

@@

  Adopting a well-known class of an epidemic model of virus infection and recovery, this study used stochastic network-based mathematical models to simulate contagion dynamics of misinformation and its corrections. More detailed methods, including model parameterization, simulation, and more detailed analyses of simulation model results, are provided in the supplementary appendix. Behavioral parameters governing the network structure and misinformation propagation dynamics were estimated from several best-available existing nationwide representative surveys concerning citizen's social network \parencite[American National Election Study 2008–2009 Panel Survey:][]{ANES2009} and exposure to fake news during 2016 presidential election \parencite{pewfakenews}. Although these data sources were not specifically designed for obtaining the impact of (mis)information diffusion dynamics within a social network but rather conveniently chosen,\footnote{  It should be acknowledged that those sources of data is suboptimal in number of regards: (a) social network data is confined to cross-sectional, egocentric network data, and (b) they do not come from the identical data generating process covering the same periods of observations (e.g., networks recoded in 2009 has little or no bearing for fake news exposure in 2016). However, research indicates the structural compositions of citizens' core discussion network are reasonably stable over time, albeit actual responses in typical surveys are subject to a number of survey-context related factors \parencite{byungkyu2017important} or changes in one's life contexts \parencite{small2015stable}. Also, since most of the prior applications concerning fake news and misinformation from observational studies typically look at representative samples, we intentionally have chosen those data since these types of data are more commonly used in the social sciences, which provide best-available up-to-date relevant information for the current application. See also \textit{A Stochastic Dynamic Model of Networks} section below and supplemental appendix for a detail.} it nevertheless provides reasonable starting point for formally model more realistic network structures, on which we later further parameterizing the assumptions and mechanisms postulated in the previous discussions of complex diffusion dynamics. All simulation models were based on the \emph{EpiModel} R software package \parencite{epimodel}.

    Utilizing simulation-based approach in modeling (mis)information propagation is not entirely new \parencite[e.g.,][]{tambuscio2015fact, jin2013epidemiological, acemoglu2010spread, zhao2010weak}. Yet these studies have typically relied on a series of deterministic compartmental models, where diffusion dynamics were derived by differential equations representing analytic epidemic systems \parencite[e.g., see][]{zhao2010weak}. While they offer valuable insights of how information spread in a given system, they do not explicitly represent contact phenomena in a network, therefore somewhat limited in realistically representing the evolving dynamic networks on which information and behaviors are spread conditional on the network structure. Our approach was to use an exponential random graph model, or ERGM framework \parencite{robins2007introduction, morris2008specification}, to simulate a series of dynamically evolving networks of different topological properties (see also below section XXX for a more detail). The ERGM framework is now regarded the most versatile yet flexible method for identifying how a given network is formed and evolves based on the underlying generative principles \parencite{cranmer2017navigating, robins2007introduction}, and for a simulation context, it offers a convenient yet powerful tool for independently manipulating topological structures of a network based on such generative mechanisms by a researcher's choice \parencite{morris2008specification, leifeld2017temporal}. In our present application, we therefore simulated a baseline model (based on available empirical evidence) and several counterfactual models (that differ in some key parameters) of network evolution models, and further submit them to different stochastic models of misinformation propagations vs. corrective message adoptions. Our focus here is therefore on the role of varying level of attitudinal homophily as a driving force of network evolutions, and how attitudinal compositions of one's local networks (as end-results of attitudinal homophily) interactively influence simple vs. complex contagion dynamics at population level in conjunction with network topologies.

    Therefore, our approach differs from prior studies that have used similar simulation-based approaches in three critical aspects: (a) instead of relying on purely hypothetical scenarios and parameter values, we augment our model by supplying more realistic parameter values based on existing empirical evidence; (b) our model explicitly takes into account the stochastic nature of network evolution and its endogenous influence over time; and finally, (c) we explicitly model how attitudinal compositions of one's immediate neighbors and local-level contagion dynamics (i.e., simple vs. complex contagions) interact to yield propagation dynamics at the population level. Arguably, this approach obviously still greatly simplifies the complexity of the real world. However, in doing so, it provides an analytical tool for identifying the effects topologies and attitudinal compositions of a network, and the resulting implications for the simple vs. complex diffusion dynamics of misinformation and corrections in a typical social network.

@@

  In order to explore role of network topologies in diffusion dynamic in evolving networks, we require to explicitly specify a network structure and how such structure evolves over time according to some underlying generative principles. To this end, we first model formation and dissolution of citizens' political discussion ties over time (through which misinformation and corrections are communicated) using separable temporal ERGMs (STERGMs), using \enquote{summary statistics} drawn from egocentric network samples \parencite{krivitsky2017inference}. That is, we estimate and simulate \emph{complete networks} that consistently reproduce observed properties of egocentrically sampled information from 2009 ANES data.\footnote{ Of course, this approach requires certain assumptions in order to define a model that represents a distribution of networks that are centered on the observed properties. See \citeauthor{krivitsky2017inference} (\citeyear{krivitsky2017inference}) for a detailed discussion of this class of models and their properties.} Those observed properties, or summary target statistics, include: density (i.e., average number of discussion partners, or \enquote{alters}), average expected number of degrees among Democrats and Republicans excluding Independents (i.e., main effect of partisanship), the extent of partisan homophily (i.e., interaction between ego and alter partisanship), and number of people who posses multiple contacts (i.e., degree distributions).\footnote{  In 2009 ANES panel, there are a total of 1258 self-identified Democrats and 1142 self-identified Republicans (including respective party leaners), excluding 336 Independents. Those 2400 partisans (i.e., Democrats plus Republicans) have on average 1.975 discussion partners. Since in 2009 ANES respondents theoretically name up to three alters, this translate into egonetwork density of 0.66 (=1.975/3). For the main effect of partisanship, the marginal distribution of number of named alters by egos did not differ by partisanship, meaning that the expected mean degree for each partisan group is proportional to observed density and the size of each group. For political homophily, data indicate that 23\% of reported discussion ties exist across Democrats and Republicans. For degree distribution statistic, we supply a range of plausible yet arbitrary values since 2009 ANES does not have summary network size measure. See supplemental appendix for a detail.} Based on those information, we establish our (a) baseline small-world with political homophily network model, which consist of 2400 nodes that comprised of 1258 Democrats and 1142 Republicans. In this model, we set approximately 70\% of all exiting ties to be politically homophilous based on observed level of political homogeneity in ANES egocentric networks. We also specify a differential relationship dissolution conditional on partisan homophily of a tie \parencite[e.g., selective "unfriending":][]{noel2011unfriending}, such that any tie within the same partisan group is expected to last longer approximately twice than cross-partisan discussion ties. 
    In addition, in order to facilitate a robust comparison of different network typologies, we additionally specify several counterfactual models of network structure: (b) Erdős–Rényi random network conditional on observed density, (c) a chain network where all nodes have no more than two connections at the same time,\footnote{  This is done by first simulating a lattice network (e.g., all nodes have exactly the same number of degrees), but due to the stochastic nature of evolving network in which ties are randomly created and dissolved, the resulting network may not have the uniform degree distribution than it would have been under strict lattice network.} and (d) a small-world network \textit{absent of partisan homophily} in tie formation and tie dissolution model. Below Figure \ref{fig:Figure1} shows the simulated cross-sectional networks from each of the network topologies, and Table \ref{tab:Table1} reports some key descriptive statistics from each of the networks.
\begin{center} -- Table \ref{tab:Table1} and Figure \ref{fig:Figure1} about here -- \end{center}

% Table 1.
\begin{table}[!htbp] \centering 
  \caption{\\ Descriptive Statistics of Cross-sectional Networks from Simulated Network Structures} 
  \label{tab:Table1} 
\begin{tabular}{@{\extracolsep{5pt}} ccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \textbf{E-R} & \textbf{Chain} &  \textbf{SW No-Homophily} & \textbf{SW Homophily} \\ 
\hline \\[-1.8ex] 
Density & 0.000272 & 0.000275 & 0.000274 & 0.000268 \\ 
Mean degree & 1.305 & 1.318 & 1.315 & 1.287 \\ 
Max degree & 10.0 & 4.0 & 10.0 & 10.0 \\ 
Clustering coefficients & 0.0 & 0.0 & 0.02 & 0.01 \\ 
Mean distance & 3.0 & 2.4 & 3.6 & 3.4 \\ 
Homophilous ties & 50.7\% & 47.3\% & 51.2\% & 74.6\% \\ 
Heterophilous ties & 49.3\% & 52.7\% & 48.8\% & 25.4\% \\ 
\hline \\[-1.8ex]  
\end{tabular} 
\begin{tablenotes}[para,flushleft]
\small \vspace{0.15in}
\textbf{Note:} E-R: Erdos Renyi random network. Chain: Chain network. SW No-Homophily: Small-world network absent of homophily. SW Homophily: Small-world network with partisan homophily. 
\end{tablenotes}
\end{table}

@@

\subsection{Misinformation Transmission and Progression}

\subsection{Corrective Message Exposure and a Recovery from Misperception}

\printbibliography
%\newpage
%\begingroup
%\parindent 0pt
%\parskip 1ex
%\def\enotesize{\normalsize}
%\theendnotes
%\endgroup
\end{document}

>>>>

% declare documents and packages
\documentclass[man, 12pt, a4paper, nolmodern, noextraspace]{apa6}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[american]{babel}
\usepackage{csquotes}
\usepackage{amsmath}
\usepackage{microtype}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{url}
\usepackage{times}
\renewcommand{\footnotesize}{\small} % use 11pt for footnotes
\usepackage[doublespacing]{setspace} % single space for footnotes
%\setlength{\skip\footins}{1.5pc plus 1pt} % some more space between footnote and main body
% this command put all footnotes to endnote at the end of the paper
\usepackage{endnotes}
\let\footnote\endnote

@@

\note{Word count: 5183 words \\ %\quickwordcount \\
    Draft date: \today \\
      \addvspace{.25in} 
      \textbf{Draft in progress. Please do not cite without permission.} \\
      Please direct any questions and inquiries to \href{mailto:hyunjin.song@univie.ac.at}{hyunjin.song@univie.ac.at}}

\begin{document}
\setcounter{page}{0}
\maketitle
  Citizens across the worlds are experiencing major changes in their news environments with the development of digital media. One of the most dramatic changes in the news environment in recent decades involves the role social networking sites (SNS) such as Facebook and Twitter play as a primary source of news outlets. Not only citizens' news consumptions are increasingly driven by such online sources \parencite{shearer2017news}, but it also appears that citizens themselves are actively participating in news dissemination on those platforms by sharing news with their peers \parencite[e.g.,][]{shearer2017news, lee2017people}.

@@

      Another line of studies based on a dual process theory of human cognitive processing and memory suggests that attitudinally-congruent misinformation tends to form a coherent mental model based on one's partisan schema and stereotypes \parencite[e.g.,][]{garrett2013undermining, Lewandowsky_2012PSPI}. Therefore, people tend to fill any gaps caused by corrections (that invalidate some parts of the existing mental model) with flawed but attitudinally congruent misinformation that is still readily accessible in their memory \parencite{Lewandowsky_2012PSPI}. Studies also find that this effect is much more likely when correction messages do not update the initial mental model that justifies misinformation \parencite{Chan_debunking_meta_2017}, when the perceived veracity of initial misinformation is high \parencite[due to fluency bias in one's cognitive processing:][]{Lewandowsky_2012PSPI}, or when individuals can generate counter-arguing reasons in support for initial misperceptions \parencite{garrett2013undermining, Chan_debunking_meta_2017}. Most importantly, due to limitations of strategic memory search processes (i.e., it requires more effortful processing), people may still rely on negated misinformation in subsequent reasoning \emph{even when they remember such information is factually incorrect} \parencite{Lewandowsky_2012PSPI}. Therefore, even in the face of seemingly effective corrections, the effect of misperceptions lingers and continue to exert influence \parencite{thorson_2016}.

  Under certain situations, it seems that citizens \emph{indeed} can adhere factual information based on correction messages despite their perpetual partisan bias \parencite[e.g.,][]{Wood2018, nyhan2017taking}. Yet as \citeauthor{margolin2017} (\citeyear{margolin2017}) note, it appears that such effects often require special \emph{social context}. This observation is indeed much warranted, as most of the previous studies concerning misinformation and the effect of fact-checking messages are conducted in an experimental context with a single-shot, \emph{asocial} correction message from media professionals and fact-checking organizations \parencite[e.g.,][]{nyhan2010corrections,garrett2013undermining,weeks2015emotions}. There is also a suggestive evidence that fact-checking and corrective messages from one's peers in their social networks -- what we would call a \enquote{social correction} -- are more likely to be effective in reducing misperceptions \parencite[e.g.,][]{margolin2017, bode2017see}. In what follow, we review several theoretical accounts of such \emph{socially-based} correction messages on misinformation and fake news.

@@

  If the diffusion of (mis)information and adoption of corrective messages may non-trivially dependent upon a structure of a given network and its attitudinal composition, then how a typical (online) social network is structured in terms of its topological features, and how pervasive is \enquote{attitudinal} homophily, as \emph{the} critical factor determining attitudinal compositions of one's social network? How such structural features affect the overall diffusion dynamics empirically? This is indeed important questions to ask, yet it is surprisingly difficult to establish convincing evidence of the impact of such factors in complex contagion dynamics using purely observational and experimental approaches.

@@

    Similarly to attitudinal composition of citizen's social networks, the question of how exactly the structure of citizens' social networks exhibits certain topological properties have attracted considerable interests among scholars. Prior observations on this topic suggests that (online) social networks tend to exhibit \enquote{small-world} like properties \parencite{kumar2010structure, ugander2011anatomy}, where most nodes can be reached from every other nodes by a small number of steps. This is due to the fact that a small minority of \enquote{hub} nodes posses a disproportionate number of links, providing a global bridge between smaller, more strongly interconnected local clusters \parencite{barabasi2004linked}. This structure has traditionally been known for information to be spread globally through such hubs \parencite[e.g.,][]{Bakshy_2012}, following \citeauthor{granovetter1977strength}'s (\citeyear{granovetter1977strength}) seminal account of strength of weak ties. However, emerging evidence suggests that such a small-world structure may hinder a given diffusion process to globally saturate the network, since such hubs can serve as a \enquote{bottleneck} of a contagion process, therefore diffusions are stay localized and compartmentalized \parencite[e.g.,][]{gonzalez2017decoding, centola2007complex, zhao2010weak}. Similarly, while the existence of tightly-knit neighborhoods in a typical small-world network supports a complex-type behavioral propagation \parencite[i.e., an innovation that runs counter to prevalent norms and values:][]{Centola2010Sience}, the existence of hubs may not favor complex contagion to be globally propagated into a given network. This is due to the fact that such hubs require a much higher number of active neighbors to meet the threshold of adopting a controversial innovation (i.e., complex contagions), and even if such hubs are activated, they alone cannot further activate their immediate neighbors to adopt such  innovations in the absence of other local nodes that provide multiple independent reinforcements \parencite{Centola2007449, zhao2010weak, gonzalez2017decoding}.

    Therefore, whether topological properties of a given network actually hinder or facilitate a contagion process may critically dependent upon specific thresholds values for complex contagions and its interaction with the structure of network \parencite[e.g.,][]{Centola2007449}, which is often hard to be tractable empirically. Moreover, especially for the case of misinformation and its corrections, a simple contagion and a complex contagion may simultaneously affect a local-level attitudinal composition of one's neighbors \parencite[i.e., those who endorse and believe misinformation vs. those who do not: e.g.,][]{campbell2013complex}. Last but not least, all of aforementioned factors are endogenously determined over time in evolving dynamic network. As a consequence, a proper identification regarding the exact impacts of its attitudinal composition and structural properties of a given network often necessitates systematic comparisons between the observed network and a number of plausible counterfactuals. Doing so further requires (a) the ability to compared different thresholds values for complex contagions, and (b) the ability to independently manipulate topological structures of a social network \parencite[e.g.,][]{Centola2010Sience, gonzalez2017decoding}. For this reason, a rigorous empirical test of such factors is often impossible for observational studies despite its scientific and practical importance.

@@

  Adopting a well-known class of an epidemic model of virus infection and recovery, this study used stochastic network-based mathematical models to simulate contagion dynamics of misinformation and its corrections. More detailed methods, including model parameterization, simulation, and more detailed analyses of simulation model results, are provided in the supplementary appendix. Behavioral parameters governing the network structure and misinformation propagation dynamics were estimated from several best-available existing nationwide representative surveys concerning citizen's social networks \parencite[American National Election Study 2008–2009 Panel Survey:][]{ANES2009} and exposure to fake news during 2016 presidential election \parencite{pewfakenews,guess2018selective,allcott2017social}. Although these data sources were not specifically designed for obtaining the impact of (mis)information diffusion dynamics within a social network, it nevertheless provides a reasonable starting point for formally modeling more realistic network structures, on which we later further parameterizing the assumptions and mechanisms postulated in the previous discussions of complex diffusion dynamics.\footnote{ It should be acknowledged that those sources of data is suboptimal in number of regards: (a) social network data is confined to cross-sectional, egocentric network data, and (b) they do not come from the identical data generating process covering the same periods of observations (e.g., networks recoded in 2009 has little or no bearing for fake news exposure in 2016). However, research indicates the structural compositions of citizens' core discussion network are reasonably stable over time, albeit actual responses in typical surveys are subject to a number of survey-context related factors \parencite{byungkyu2017important} or changes in one's life contexts \parencite{small2015stable}. Also, since most of the prior applications concerning fake news and misinformation from observational studies typically look at representative samples, we intentionally have chosen those data since these types of data are more commonly used in the social sciences, and provide best-available up-to-date relevant information for the current application. See also \textit{A Stochastic Dynamic Model of Networks} section below and supplemental appendix for a detail.} All simulation models were based on the \emph{EpiModel} R software package \parencite{epimodel}.

    Utilizing simulation-based approach in modeling (mis)information propagation is not entirely new \parencite[e.g.,][]{tambuscio2015fact, jin2013epidemiological, zhao2010weak}. Yet these studies have typically relied on a series of deterministic compartmental models, where diffusion dynamics were derived by differential equations representing analytic epidemic systems \parencite[e.g., see][]{zhao2010weak}. While they offer valuable insights of how information spread in a given system, they do not explicitly represent contact phenomena in a network, therefore somewhat limited in realistically representing the evolving dynamic networks on which information and behaviors are spread conditional on the network structure. Our approach was to use an exponential random graph model (ERGM) framework \parencite{robins2007introduction, morris2008specification} to simulate a series of dynamically evolving networks of different topological properties (see also below section \textit{A Stochastic Dynamic Model of Networks} for a more detail). The ERGM framework is now regarded the most versatile yet flexible method for identifying how a given network is formed and evolves based on the underlying generative principles \parencite{cranmer2017navigating, robins2007introduction}, and for a simulation context, it offers a convenient yet powerful tool for independently manipulating topological structures of a evolving network based on such generative mechanisms by a researcher's choice \parencite{morris2008specification, leifeld2017temporal}. In our present application, we simulated a baseline network evolution model based on available empirical evidence along with several counterfactual models that differ in their structural characteristics, and further submit them to different stochastic models of misinformation propagations vs. corrective message adoptions. Our focus here is therefore on the role of varying network topologies and level of attitudinal homophily as a driving force of network evolutions, and how attitudinal compositions of one's local networks (as end-results of attitudinal homophily) interactively influence simple vs. complex contagion dynamics at population level in conjunction with network topologies.

    Therefore, our approach differs from prior studies that have used similar simulation-based approaches in three critical aspects: (a) instead of relying on purely hypothetical scenarios and parameter values, we augment our model by supplying more realistic parameter values, where available, based on existing empirical evidence; (b) our model explicitly takes into account the stochastic nature of network evolution and its endogenous influence over time; and finally, (c) we explicitly model how attitudinal compositions of one's immediate neighbors and local-level contagion dynamics (i.e., simple vs. complex contagions) interact to yield propagation dynamics at the population level. Arguably, this approach obviously still greatly simplifies the complexity of the real world. However, in doing so, it provides an analytical tool for identifying the effects topologies and attitudinal compositions of a network, and the resulting implications for the simple vs. complex diffusion dynamics of misinformation and corrections in a typical social network.

@@

  In order to explore role of network topologies in diffusion dynamic in evolving networks, we require to explicitly specify a generative principle of network structure and how such structure evolves over time. To this end, we first model formation and dissolution of citizens' political discussion ties over time through which misinformation and corrections are communicated. Using separable temporal ERGMs \parencite{krivitsky2017inference}, we estimate and simulate \emph{complete networks} that consistently reproduce observed properties of \enquote{summary statistics} drawn from egocentric network from 2009 ANES data.\footnote{  Of course, this approach requires certain assumptions in order to define a model that represents a distribution of networks that are centered on the observed properties. See \citeauthor{krivitsky2017inference} (\citeyear{krivitsky2017inference}) for a detailed discussion of this class of models and their properties.} Those observed properties, or summary target statistics, include: density (i.e., average number of discussion partners, or \enquote{alters}), average expected number of degrees among Democrats and Republicans (i.e., main effect of partisanship, excluding Independents), the extent of partisan homophily (i.e., interaction between ego and alter partisanship), and number of people who posses multiple contacts (i.e., degree distributions).\footnote{\label{note6}  In 2009 ANES panel, there are a total of 1258 self-identified Democrats and 1142 self-identified Republicans (including respective party leaners), excluding 336 Independents. Those 2400 partisans (i.e., Democrats plus Republicans) have on average 1.975 discussion partners. Since in 2009 ANES respondents theoretically name up to three alters, this translate into egonetwork density of 0.66 (=1.975/3). For the main effect of partisanship, the marginal distribution of number of named alters by egos did not differ by partisanship, meaning that the expected mean degree for each partisan group is proportional to observed density and the size of each group. For political homophily, data indicate that 23\% of reported discussion ties exist across Democrats and Republicans. For degree distribution statistic, we supply a range of plausible yet arbitrary values since 2009 ANES does not have summary network size measure. See supplemental appendix for a detail.} Based on those information, we establish our (a) baseline small-world with political homophily network model, which consist of 2400 nodes that comprised of 1258 Democrats and 1142 Republicans. In this model, we set approximately 70\% of all exiting ties to be politically homophilous based on observed level of political homogeneity in ANES egocentric networks. We also specify a differential relationship dissolution conditional on partisan homophily of a tie \parencite[e.g., selective \enquote{unfriending}:][]{noel2011unfriending}, such that any tie within the same partisan group is expected to last longer approximately twice than cross-partisan discussion ties. 
    In addition, in order to facilitate a robust comparison of different network typologies, we additionally specify several counterfactual models of network structure: (b) Erdős–Rényi random network conditional on observed density, (c) a chain network where all nodes have no more than two connections at the same time,\footnote{  This is done by first simulating a lattice network (e.g., all nodes have exactly the same number of degrees), but due to the stochastic nature of evolving network in which ties are randomly created and dissolved, the resulting network may not have the uniform degree distribution than it would have been under strict lattice network.} and (d) a small-world network \textit{absent of partisan homophily} in tie formation and tie dissolution model. Below Figure \ref{fig:Figure1} shows the simulated cross-sectional views of networks, and Table \ref{tab:Table1} reports key descriptive statistics for each of the networks in terms of their topological properties.
\centerline{ -- Table \ref{tab:Table1} and Figure \ref{fig:Figure1} about here -- }

% Table 1.
\begin{table}[!htbp] \centering 
  \caption{\\ Descriptive Statistics of Cross-sectional Networks from Simulated Network Structures} 
  \label{tab:Table1} 
\begin{tabular}{@{\extracolsep{5pt}} ccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \textbf{E-R} & \textbf{Chain} &  \textbf{SW No-Homophily} & \textbf{SW Homophily} \\ 
\hline \\[-1.8ex] 
Graph density & 0.000269 & 0.000268 & 0.000277 & 0.000272 \\ 
Mean degree & 1.292 & 1.287 & 1.327 & 1.303 \\ 
Max degree & 10.0 & 4.0 & 12.0 & 12.0 \\ 
Clustering coefficients & 0.0 & 0.0 & 0.02 & 0.01 \\ 
Mean distance & 2.2 & 2.2 & 4.5 & 3.4 \\ 
Homophilous ties & 49.4\% & 50.5\% & 52.3\% & 68.5\% \\ 
Heterophilous ties & 50.6\% & 49.5\% & 47.7\% & 31.5\% \\ 
\hline \\[-1.8ex]  
\end{tabular} 
\begin{tablenotes}[para,flushleft]
\small \vspace{0.15in}
\textbf{Note:} E-R: Erdos Renyi random network. Chain: Chain network. SW No-Homophily: Small-world network absent of homophily. SW Homophily: Small-world network with partisan homophily. 
\end{tablenotes}
\end{table}

@@

  Adapting a well-known class of epidemic models, we simulated the dynamics of misinformation exposure and its correction on top of those stochastic network evolution models. We setup an environment consist of 2400 partisan actors (based on 2009 ANES panel survey's partisan composition: see footnote \ref{note6} for a detail) and simulate the diffusion dynamcics over 1000 simulated hours (roughly corresponds to 41 days) with 100 replicated simulations per each network topologies. We fit a series of SEIR (\enquote{suseptable - exposed - infected - recovered}) dynamic models for (a) a simple contagion of misinformation (i.e., a progression from \enquote{suspected} to \enquote{exposed}), (b) a progression from \enquote{exposed} to \enquote{infected} (i.e., those who believe a given misinformation conditional on exposure), and (c) a progression from \enquote{infected} to \enquote{recovered} conditional on exposure and adoption of corrective messages from their peers.

\subsection{Misinformation Transmission}

    To date, how many times people exactly \enquote{share} fake news and misinformation on their social networks is not yet known to public. However, a recent study by \citeauthor{allcott2017social} (\citeyear{allcott2017social}) report that, for all 65 fake news websites they have identified during 2016 election, an individual visited such sites 0.64 times per month on average, and among them 41.8\% of all visits originate from social media traffics. Similarly, \citeauthor{guess2018selective} (\citeyear{guess2018selective}) report that of 27.4\% inidivduals who visited fake news websites, 22.1\% of all visites originate from social media.

  We start by setting our baseline prevalence, of which we arbitrary set 15\% of partisans are initially affected by misinformation and fake news. We then set our baseline \enquote{act rate} (i.e., the average number of transmissible acts per unit time per discussion tie) by taking the average of estimates from \citeauthor{allcott2017social} (\citeyear{allcott2017social}) and from \citeauthor{guess2018selective} (\citeyear{guess2018selective}), which yields 0.164 times per hour per discussion tie.\footnote{ \citeauthor{allcott2017social} report 0.64 $\times$ 41.8\% = 0.268 shares per unit time, while \citeauthor{guess2018selective} report 0.274 $\times$ 22.1\% = 0.060 shares per unit time. This averages to 0.164 times per unit time. We additionally assume all fake news website visits are driven by information sharing on social media.} We further assume 15\% of those who initially affected are only Republicans, given that the vast majority of identified fake news articles during 2016 U.S. presidential election were overwhelmingly pro-Trump \parencite[e.g.,][]{guess2018selective}. Here, we assume misinformation always flows from the \enquote{infected} (i.e., those who already believe given misinformation) to their suseptable peers (i.e., those who are not yet exposed to misinformation), but those who are being mere exposed do not contribute to misinformaiton flow (since the mere exposure does not necessarily mean they are endorsing such information) nor any other sources of exposure to misinformation is possible (i.e., partisan media exposure) for an individual to encounter such misinformation.
